Manuscript 1 code

### 2024-11-20 manuscript_1_code.Rmd

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library(here)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(vegan)
library(glmmTMB)
library(car)
library(effects)
library(DHARMa)
library(stringr)
library(MuMIn)
library(codyn)
library(nlme)
library(mgcv)
library(emmeans)
library(dplyr)
library(janitor)
library(knitr)
```

@@

cover_mod <- glmmTMB(cover_trans ~ richness*habitat + (1|site/location) + (1|year), family = beta_family(), data = alpha_diversity_quad_macro)
summary(cover_mod)
Anova(cover_mod)
#                     Chisq Df Pr(>Chisq)    
# richness         30916.73  1  < 2.2e-16 ***
# habitat            141.78  3  < 2.2e-16 ***
# richness:habitat   294.54  3  < 2.2e-16 ***
hist(residuals(cover_mod)) # looks fine
plot(residuals(cover_mod) ~ fitted(cover_mod)) # negative trend but it can't be less than 0 so not too concerned. some wave pattern
r.squaredGLMM(cover_mod) # 0.631, 0.650
emtrends(cover_mod, pairwise ~ habitat, var = "richness") # backreef not different from forereef, forereef also not different
# contrast                    estimate     SE  df z.ratio p.value
# Fringing - Backreef          0.34720 0.0241 Inf  14.409  <.0001
# Fringing - Forereef 10m      0.34865 0.0240 Inf  14.514  <.0001
# Fringing - Forereef 17m      0.38953 0.0250 Inf  15.599  <.0001
# Backreef - Forereef 10m      0.00144 0.0205 Inf   0.071  0.9999
# Backreef - Forereef 17m      0.04232 0.0214 Inf   1.979  0.1957
# Forereef 10m - Forereef 17m  0.04088 0.0209 Inf   1.952  0.2063
```

@@

r.squaredGLMM(cover_mod_fg) # 0.64, 0.65
emtrends(cover_mod_fg, pairwise ~ habitat, var = "functional_richness") 
# contrast                    estimate     SE  df z.ratio p.value
# Fringing - Backreef            0.214 0.0266 Inf   8.051  <.0001
# Fringing - Forereef 10m        0.346 0.0261 Inf  13.252  <.0001
# Fringing - Forereef 17m        0.530 0.0262 Inf  20.244  <.0001
# Backreef - Forereef 10m        0.132 0.0231 Inf   5.708  <.0001
# Backreef - Forereef 17m        0.316 0.0232 Inf  13.617  <.0001
# Forereef 10m - Forereef 17m    0.184 0.0223 Inf   8.241  <.0001

@@

r.squaredGLMM(synch_cov_mod) # error??
emtrends(synch_cov_mod, pairwise ~ habitat, var = "synchrony") # forereef 17 distinct from all other habitats

@@

### model at site level
```{r site richness mod}
site_cover_mod <- glmmTMB(cover_trans ~ richness*habitat + (1|site_habitat) + (1|year), family = beta_family(), data = alpha_diversity_site_macro)
summary(site_cover_mod)
Anova(site_cover_mod)
#                    Chisq Df Pr(>Chisq)    
# richness         58.9195  1  1.643e-14 ***
# habitat           6.6247  3    0.08487 .  
# richness:habitat  3.2967  3    0.34811    
hist(residuals(site_cover_mod)) # some positive skew, not the worst thing ever
plot(residuals(site_cover_mod) ~ fitted(site_cover_mod)) # a little heteroscedastic but can't be less than 0
r.squaredGLMM(site_cover_mod) # 0.2212534, 0.5990825
emtrends(site_cover_mod, pairwise ~ habitat, var = "richness") # no difference between the habitats
# contrast                    estimate     SE  df z.ratio p.value
# Fringing - Backreef          -0.0163 0.0394 Inf  -0.413  0.9762
# Fringing - Forereef 10m       0.0639 0.0503 Inf   1.271  0.5818
# Fringing - Forereef 17m       0.0334 0.0612 Inf   0.546  0.9476
# Backreef - Forereef 10m       0.0801 0.0460 Inf   1.742  0.3020
# Backreef - Forereef 17m       0.0497 0.0583 Inf   0.853  0.8289
# Forereef 10m - Forereef 17m  -0.0304 0.0651 Inf  -0.467  0.9662
```

@@

hist(residuals(site_cover_mod_fg)) # positive skew
plot(residuals(site_cover_mod_fg) ~ fitted(site_cover_mod_fg)) # a little heteroscedastic but can't be less than 0
r.squaredGLMM(site_cover_mod_fg) # 0.180, 0.563
emtrends(site_cover_mod_fg, pairwise ~ habitat, var = "functional_richness") # no difference between the habitats
# Fringing - Backreef           0.0379 0.0825 Inf   0.459  0.9679
# Fringing - Forereef 10m       0.1846 0.0930 Inf   1.985  0.1935
# Fringing - Forereef 17m       0.0933 0.0960 Inf   0.972  0.7655
# Backreef - Forereef 10m       0.1467 0.0986 Inf   1.488  0.4447
# Backreef - Forereef 17m       0.0554 0.1018 Inf   0.545  0.9480
# Forereef 10m - Forereef 17m  -0.0913 0.1093 Inf  -0.835  0.8377

@@

```{r synchrony cover mod}
synch_cov_mod_site <- glmmTMB(cover_mean ~ synchrony*habitat, family = beta_family(), data = diversity_stability_synchrony_site)
summary(synch_cov_mod_site)
Anova(synch_cov_mod_site)
#                    Chisq Df Pr(>Chisq)  
# synchrony         2.4184  1    0.11992  
# habitat           7.8002  3    0.05033 .
# synchrony:habitat 8.8866  3    0.03084 *
hist(residuals(synch_cov_mod_site)) # slight skew
plot(residuals(synch_cov_mod_site) ~ fitted(synch_cov_mod_site)) # looks fine
emtrends(synch_cov_mod_site, pairwise ~ habitat, var = "synchrony") # forereef 17 different from backreef
# contrast                    estimate   SE  df z.ratio p.value
# Fringing - Backreef           -4.299 3.38 Inf  -1.270  0.5821
# Fringing - Forereef 10m       -0.279 3.22 Inf  -0.087  0.9998
# Fringing - Forereef 17m        3.675 3.78 Inf   0.973  0.7649
# Backreef - Forereef 10m        4.019 1.98 Inf   2.031  0.1764
# Backreef - Forereef 17m        7.973 2.80 Inf   2.848  0.0228
# Forereef 10m - Forereef 17m    3.954 2.59 Inf   1.526  0.4217
```

@@

```{r richness stability mod}
rich_stab_mod <- glmmTMB(cover_stability~ richness_mean*habitat + (1|site), family = Gamma(link = "log"), data = diversity_stability_synchrony)
summary(rich_stab_mod)
Anova(rich_stab_mod)
#                          Chisq Df Pr(>Chisq)    
# richness_mean         2721.438  1  < 2.2e-16 ***
# habitat                325.876  3  < 2.2e-16 ***
# richness_mean:habitat   33.926  3  2.054e-07 ***
hist(residuals(rich_stab_mod)) # looks good
plot(residuals(rich_stab_mod) ~ fitted(rich_stab_mod)) # heteroskedastic - bring to group
r.squaredGLMM(rich_stab_mod) # 0.803, 0.808
emtrends(rich_stab_mod, pairwise ~ habitat, var = "richness_mean") # fringing not diff from outer17, backreef not different from outer10
# contrast                     estimate     SE  df z.ratio p.value
# Fringing - Backreef          0.186670 0.0414 Inf   4.512  <.0001
# Fringing - Forereef 10m      0.187196 0.0455 Inf   4.114  0.0002
# Fringing - Forereef 17m      0.027112 0.0463 Inf   0.586  0.9364
# Backreef - Forereef 10m      0.000525 0.0405 Inf   0.013  1.0000
# Backreef - Forereef 17m     -0.159559 0.0411 Inf  -3.885  0.0006
# Forereef 10m - Forereef 17m -0.160084 0.0439 Inf  -3.647  0.0015
```

@@

hist(residuals(rich_stab_mod_fg)) # looks fine
plot(residuals(rich_stab_mod_fg) ~ fitted(rich_stab_mod_fg)) # same as usual 
r.squaredGLMM(rich_stab_mod_fg) # 0.819, 0.828
emtrends(rich_stab_mod_fg, pairwise ~ habitat, var = "functional_richness_mean") 
#  fringing different from all others
# contrast                    estimate     SE  df z.ratio p.value
# Fringing - Backreef           0.1872 0.0468 Inf   3.999  0.0004
# Fringing - Forereef 10m       0.2348 0.0519 Inf   4.524  <.0001
# Fringing - Forereef 17m       0.2040 0.0489 Inf   4.173  0.0002
# Backreef - Forereef 10m       0.0477 0.0472 Inf   1.011  0.7432
# Backreef - Forereef 17m       0.0169 0.0439 Inf   0.385  0.9806
# Forereef 10m - Forereef 17m  -0.0308 0.0475 Inf  -0.648  0.9164
#
```

@@

```{r synchrony staability mod}
synch_stab_mod <- glmmTMB(cover_stability ~ synchrony*habitat + (1|site), family = Gamma("log"), data = diversity_stability_synchrony)
summary(synch_stab_mod)
Anova(synch_stab_mod)
#                     Chisq Df Pr(>Chisq)    
# synchrony         1308.33  1  < 2.2e-16 ***
# habitat            103.13  3  < 2.2e-16 ***
# synchrony:habitat  327.97  3  < 2.2e-16 ***
hist(residuals(synch_stab_mod)) # looks good
plot(residuals(synch_stab_mod) ~ fitted(synch_stab_mod)) # same issue as always
r.squaredGLMM(synch_stab_mod) # 0.63, 0.71
emtrends(synch_stab_mod, pairwise ~ habitat, var = "synchrony") # fringing and backreef not different
# contrast                    estimate    SE  df z.ratio p.value
# Fringing - Backreef            0.181 0.107 Inf   1.695  0.3262
# Fringing - Forereef 10m        1.492 0.142 Inf  10.487  <.0001
# Fringing - Forereef 17m        2.436 0.149 Inf  16.336  <.0001
# Backreef - Forereef 10m        1.311 0.157 Inf   8.346  <.0001
# Backreef - Forereef 17m        2.255 0.161 Inf  14.008  <.0001
# Forereef 10m - Forereef 17m    0.944 0.181 Inf   5.222  <.0001
```

@@

#                        Chisq Df Pr(>Chisq)    
# richness_mean         17.645  1  2.662e-05 ***
# habitat               32.229  3  4.682e-07 ***
# richness_mean:habitat  8.075  3    0.04449 *  
hist(residuals(rich_stab_mod_site)) # blocky but fine
plot(residuals(rich_stab_mod_site) ~ fitted(rich_stab_mod_site)) # heteroskedastic - bring to group
r.squaredGLMM(rich_stab_mod_site) # 0.671
emtrends(rich_stab_mod_site, pairwise ~ habitat, var = "richness_mean") # no differences across habitats
# contrast                    estimate    SE  df z.ratio p.value
# Fringing - Backreef           0.1854 0.120 Inf   1.539  0.4138
# Fringing - Forereef 10m      -0.2155 0.145 Inf  -1.484  0.4471
# Fringing - Forereef 17m      -0.2876 0.214 Inf  -1.344  0.5349
# Backreef - Forereef 10m      -0.4009 0.165 Inf  -2.432  0.0711
# Backreef - Forereef 17m      -0.4730 0.228 Inf  -2.077  0.1607
# Forereef 10m - Forereef 17m  -0.0721 0.242 Inf  -0.298  0.9908
```

@@

hist(residuals(rich_stab_mod_site_fg)) # blocky but fine
plot(residuals(rich_stab_mod_site_fg) ~ fitted(rich_stab_mod_site_fg)) # heteroskedastic
r.squaredGLMM(rich_stab_mod_site_fg) # 0.563
emtrends(rich_stab_mod_site_fg, pairwise ~ habitat, var = "functional_richness_mean") 
# Fringing different from outer 10
# contrast                    estimate    SE  df z.ratio p.value
# Fringing - Backreef           0.2552 0.254 Inf   1.005  0.7463
# Fringing - Forereef 10m       1.2611 0.469 Inf   2.686  0.0364
# Fringing - Forereef 17m       0.3543 0.287 Inf   1.235  0.6045
# Backreef - Forereef 10m       1.0059 0.492 Inf   2.044  0.1720
# Backreef - Forereef 17m       0.0991 0.323 Inf   0.307  0.9900
# Forereef 10m - Forereef 17m  -0.9068 0.510 Inf  -1.778  0.2838
```

@@

### model
```{r synchrony staability mod site}
synch_stab_mod_site <- glmmTMB(cover_stability ~ synchrony*habitat, family = Gamma("log"), data = diversity_stability_synchrony_site)
summary(synch_stab_mod_site)
Anova(synch_stab_mod_site)
#                     Chisq Df Pr(>Chisq)    
# synchrony         22.0686  1  2.631e-06 ***
# habitat            9.3333  3    0.02517 *  
# synchrony:habitat 10.5561  3    0.01439 *  
hist(residuals(synch_stab_mod_site)) # almost uniform
plot(residuals(synch_stab_mod_site) ~ fitted(synch_stab_mod_site)) # same issue as always
r.squaredGLMM(synch_stab_mod_site) # 0.678
emtrends(synch_stab_mod_site, pairwise ~ habitat, var = "synchrony") # backreef is different from forereef 17, fringing is a p = 0.05
#
# contrast                    estimate    SE  df z.ratio p.value
# Fringing - Backreef           -0.256 0.971 Inf  -0.263  0.9936
# Fringing - Forereef 10m        1.324 0.916 Inf   1.445  0.4709
# Fringing - Forereef 17m        2.930 1.142 Inf   2.567  0.0503
# Backreef - Forereef 10m        1.580 0.858 Inf   1.842  0.2534
# Backreef - Forereef 17m        3.186 1.095 Inf   2.909  0.0190
# Forereef 10m - Forereef 17m    1.606 1.047 Inf   1.534  0.4172
```

@@

### spatial synchrony
```{r model spatial synchrony}
spatial_synchrony_mod <- glmmTMB(cover_stability ~ spatial_synchrony*habitat, family = Gamma("log"), data = dss_spatial_2)
hist(residuals(spatial_synchrony_mod)) # a little skew
plot(residuals(spatial_synchrony_mod) ~ predict(spatial_synchrony_mod))
summary(spatial_synchrony_mod)
Anova(spatial_synchrony_mod)
#                             Chisq Df Pr(>Chisq)    
# spatial_synchrony         16.2914  1  5.431e-05 ***
# habitat                   19.5303  3  0.0002124 ***
# spatial_synchrony:habitat  6.7239  3  0.0812368 .  
emtrends(spatial_synchrony_mod, pairwise ~ habitat, var = "spatial_synchrony") # no sig differences but close
# contrast                    estimate    SE  df z.ratio p.value
# Fringing - Backreef            0.404 1.765 Inf   0.229  0.9958
# Fringing - Forereef 10m        1.365 1.754 Inf   0.778  0.8643
# Fringing - Forereef 17m        3.112 1.882 Inf   1.654  0.3484
# Backreef - Forereef 10m        0.961 0.857 Inf   1.121  0.6766
# Backreef - Forereef 17m        2.708 1.095 Inf   2.472  0.0644
# Forereef 10m - Forereef 17m    1.747 1.078 Inf   1.621  0.3668
```

@@

### mean synchrony
```{r mean synchrony}
mean_synchrony_mod <- glmmTMB(cover_stability ~ synchrony_mean*habitat, family = Gamma("log"), data = dss_spatial_2)
hist(residuals(mean_synchrony_mod)) # chunky
plot(residuals(mean_synchrony_mod) ~ predict(mean_synchrony_mod)) # fine
summary(mean_synchrony_mod)
Anova(mean_synchrony_mod)
#                         Chisq Df Pr(>Chisq)    
# synchrony_mean         32.914  1  9.634e-09 ***
# habitat                18.523  3  0.0003430 ***
# synchrony_mean:habitat 19.062  3  0.0002654 ***
emtrends(mean_synchrony_mod, pairwise ~ habitat, var = "synchrony_mean") # no sig differences but close
# contrast                    estimate    SE  df z.ratio p.value
# Fringing - Backreef            -3.26 1.503 Inf  -2.166  0.1327
# Fringing - Forereef 10m         2.22 0.995 Inf   2.231  0.1149
# Fringing - Forereef 17m         3.37 1.220 Inf   2.764  0.0291
# Backreef - Forereef 10m         5.48 1.622 Inf   3.377  0.0041
# Backreef - Forereef 17m         6.63 1.769 Inf   3.748  0.0010
# Forereef 10m - Forereef 17m     1.15 1.363 Inf   0.846  0.8322
```

@@

### mean stability plot
```{r alpha gamma stability mod plot}
alpha_gamma_stab_mod <- glmmTMB(cover_stability ~ stability_mean*habitat, family = Gamma("log"), data = dss_spatial_2)
summary(alpha_gamma_stab_mod)
Anova(alpha_gamma_stab_mod)
#                          Chisq Df Pr(>Chisq)    
# stability_mean         61.2725  1   4.97e-15 ***
# habitat                12.9323  3   0.004785 ** 
# stability_mean:habitat  9.4832  3   0.023511 *

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library(here)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(vegan)
library(glmmTMB)
library(car)
library(effects)
library(DHARMa)
library(stringr)
library(MuMIn)
library(codyn)
library(nlme)
library(mgcv)
library(emmeans)
library(dplyr)
library(janitor)
library(knitr)
library(performance)
```

@@

cover_mod <- glmmTMB(cover_trans ~ richness*habitat + (1|site/location) + (1|year), family = beta_family(), data = alpha_diversity_quad_macro)
summary(cover_mod)
Anova(cover_mod)
#                     Chisq Df Pr(>Chisq)    
# richness         30916.73  1  < 2.2e-16 ***
# habitat            141.78  3  < 2.2e-16 ***
# richness:habitat   294.54  3  < 2.2e-16 ***
hist(residuals(cover_mod)) # looks fine
plot(residuals(cover_mod) ~ fitted(cover_mod)) # negative trend but it can't be less than 0 so not too concerned. some wave pattern
r.squaredGLMM(cover_mod) # 0.631, 0.650
performance::r2(cover_mod) # 0.634, 0.654
emtrends(cover_mod, pairwise ~ habitat, var = "richness") # backreef not different from forereef, forereef also not different
# contrast                    estimate     SE  df z.ratio p.value
# Fringing - Backreef          0.34720 0.0241 Inf  14.409  <.0001
# Fringing - Forereef 10m      0.34865 0.0240 Inf  14.514  <.0001
# Fringing - Forereef 17m      0.38953 0.0250 Inf  15.599  <.0001
# Backreef - Forereef 10m      0.00144 0.0205 Inf   0.071  0.9999
# Backreef - Forereef 17m      0.04232 0.0214 Inf   1.979  0.1957
# Forereef 10m - Forereef 17m  0.04088 0.0209 Inf   1.952  0.2063
```

@@

r.squaredGLMM(cover_mod_fg) # 0.64, 0.65
performance::r2(cover_mod_fg) # 0.63, 0.65
emtrends(cover_mod_fg, pairwise ~ habitat, var = "functional_richness") 
# contrast                    estimate     SE  df z.ratio p.value
# Fringing - Backreef            0.214 0.0266 Inf   8.051  <.0001
# Fringing - Forereef 10m        0.346 0.0261 Inf  13.252  <.0001
# Fringing - Forereef 17m        0.530 0.0262 Inf  20.244  <.0001
# Backreef - Forereef 10m        0.132 0.0231 Inf   5.708  <.0001
# Backreef - Forereef 17m        0.316 0.0232 Inf  13.617  <.0001
# Forereef 10m - Forereef 17m    0.184 0.0223 Inf   8.241  <.0001

@@

# can't compute r2 - manually fit null model. Null model won't fit without sv_trans() for some reason
null_synch_mod <- glmmTMB(sv_trans(cover_mean) ~ 1, family = beta_family(), data = diversity_stability_synchrony)
performance::r2(model = synch_cov_mod, null_model = null_synch_mod)
emtrends(synch_cov_mod, pairwise ~ habitat, var = "synchrony") # forereef 17 distinct from all other habitats

@@

### model at site level
```{r site richness mod}
site_cover_mod <- glmmTMB(cover_trans ~ richness*habitat + (1|site_habitat) + (1|year), family = beta_family(), data = alpha_diversity_site_macro)
summary(site_cover_mod)
Anova(site_cover_mod)
#                    Chisq Df Pr(>Chisq)    
# richness         58.9195  1  1.643e-14 ***
# habitat           6.6247  3    0.08487 .  
# richness:habitat  3.2967  3    0.34811    
hist(residuals(site_cover_mod)) # some positive skew, not the worst thing ever
plot(residuals(site_cover_mod) ~ fitted(site_cover_mod)) # a little heteroscedastic but can't be less than 0
r.squaredGLMM(site_cover_mod) # 0.2212534, 0.5990825
performance::r2(site_cover_mod) # 0.22, 0.59
emtrends(site_cover_mod, pairwise ~ habitat, var = "richness") # no difference between the habitats
# contrast                    estimate     SE  df z.ratio p.value
# Fringing - Backreef          -0.0163 0.0394 Inf  -0.413  0.9762
# Fringing - Forereef 10m       0.0639 0.0503 Inf   1.271  0.5818
# Fringing - Forereef 17m       0.0334 0.0612 Inf   0.546  0.9476
# Backreef - Forereef 10m       0.0801 0.0460 Inf   1.742  0.3020
# Backreef - Forereef 17m       0.0497 0.0583 Inf   0.853  0.8289
# Forereef 10m - Forereef 17m  -0.0304 0.0651 Inf  -0.467  0.9662
```

@@

hist(residuals(site_cover_mod_fg)) # positive skew
plot(residuals(site_cover_mod_fg) ~ fitted(site_cover_mod_fg)) # a little heteroscedastic but can't be less than 0
r.squaredGLMM(site_cover_mod_fg) # 0.180, 0.563
performance::r2(site_cover_mod_fg) # 0.180, 0.563
emtrends(site_cover_mod_fg, pairwise ~ habitat, var = "functional_richness") # no difference between the habitats
# Fringing - Backreef           0.0379 0.0825 Inf   0.459  0.9679
# Fringing - Forereef 10m       0.1846 0.0930 Inf   1.985  0.1935
# Fringing - Forereef 17m       0.0933 0.0960 Inf   0.972  0.7655
# Backreef - Forereef 10m       0.1467 0.0986 Inf   1.488  0.4447
# Backreef - Forereef 17m       0.0554 0.1018 Inf   0.545  0.9480
# Forereef 10m - Forereef 17m  -0.0913 0.1093 Inf  -0.835  0.8377

@@

```{r synchrony cover mod}
synch_cov_mod_site <- glmmTMB(cover_mean ~ synchrony*habitat, family = beta_family(), data = diversity_stability_synchrony_site)
summary(synch_cov_mod_site)
Anova(synch_cov_mod_site)
#                    Chisq Df Pr(>Chisq)  
# synchrony         2.4184  1    0.11992  
# habitat           7.8002  3    0.05033 .
# synchrony:habitat 8.8866  3    0.03084 *
performance::r2(synch_cov_mod_site)
hist(residuals(synch_cov_mod_site)) # slight skew
plot(residuals(synch_cov_mod_site) ~ fitted(synch_cov_mod_site)) # looks fine
emtrends(synch_cov_mod_site, pairwise ~ habitat, var = "synchrony") # forereef 17 different from backreef
# contrast                    estimate   SE  df z.ratio p.value
# Fringing - Backreef           -4.299 3.38 Inf  -1.270  0.5821
# Fringing - Forereef 10m       -0.279 3.22 Inf  -0.087  0.9998
# Fringing - Forereef 17m        3.675 3.78 Inf   0.973  0.7649
# Backreef - Forereef 10m        4.019 1.98 Inf   2.031  0.1764
# Backreef - Forereef 17m        7.973 2.80 Inf   2.848  0.0228
# Forereef 10m - Forereef 17m    3.954 2.59 Inf   1.526  0.4217
```

@@

```{r richness stability mod}
rich_stab_mod <- glmmTMB(cover_stability~ richness_mean*habitat + (1|site), family = Gamma(link = "log"), data = diversity_stability_synchrony)
summary(rich_stab_mod)
Anova(rich_stab_mod)
#                          Chisq Df Pr(>Chisq)    
# richness_mean         2721.438  1  < 2.2e-16 ***
# habitat                325.876  3  < 2.2e-16 ***
# richness_mean:habitat   33.926  3  2.054e-07 ***
hist(residuals(rich_stab_mod)) # looks good
plot(residuals(rich_stab_mod) ~ fitted(rich_stab_mod)) # heteroskedastic - bring to group
r.squaredGLMM(rich_stab_mod) # 0.803, 0.808
performance::r2(rich_stab_mod) # 0.812, 0.816
emtrends(rich_stab_mod, pairwise ~ habitat, var = "richness_mean") # fringing not diff from outer17, backreef not different from outer10
# contrast                     estimate     SE  df z.ratio p.value
# Fringing - Backreef          0.186670 0.0414 Inf   4.512  <.0001
# Fringing - Forereef 10m      0.187196 0.0455 Inf   4.114  0.0002
# Fringing - Forereef 17m      0.027112 0.0463 Inf   0.586  0.9364
# Backreef - Forereef 10m      0.000525 0.0405 Inf   0.013  1.0000
# Backreef - Forereef 17m     -0.159559 0.0411 Inf  -3.885  0.0006
# Forereef 10m - Forereef 17m -0.160084 0.0439 Inf  -3.647  0.0015
```

@@

hist(residuals(rich_stab_mod_fg)) # looks fine
plot(residuals(rich_stab_mod_fg) ~ fitted(rich_stab_mod_fg)) # same as usual 
r.squaredGLMM(rich_stab_mod_fg) # 0.819, 0.828
performance::r2(rich_stab_mod_fg)
emtrends(rich_stab_mod_fg, pairwise ~ habitat, var = "functional_richness_mean") 
#  fringing different from all others
# contrast                    estimate     SE  df z.ratio p.value
# Fringing - Backreef           0.1872 0.0468 Inf   3.999  0.0004
# Fringing - Forereef 10m       0.2348 0.0519 Inf   4.524  <.0001
# Fringing - Forereef 17m       0.2040 0.0489 Inf   4.173  0.0002
# Backreef - Forereef 10m       0.0477 0.0472 Inf   1.011  0.7432
# Backreef - Forereef 17m       0.0169 0.0439 Inf   0.385  0.9806
# Forereef 10m - Forereef 17m  -0.0308 0.0475 Inf  -0.648  0.9164
#
```

@@

```{r synchrony staability mod}
synch_stab_mod <- glmmTMB(cover_stability ~ synchrony*habitat + (1|site), family = Gamma("log"), data = diversity_stability_synchrony)
summary(synch_stab_mod)
Anova(synch_stab_mod)
#                     Chisq Df Pr(>Chisq)    
# synchrony         1308.33  1  < 2.2e-16 ***
# habitat            103.13  3  < 2.2e-16 ***
# synchrony:habitat  327.97  3  < 2.2e-16 ***
hist(residuals(synch_stab_mod)) # looks good
plot(residuals(synch_stab_mod) ~ fitted(synch_stab_mod)) # same issue as always
r.squaredGLMM(synch_stab_mod) # 0.63, 0.71
performance::r2(synch_stab_mod)
emtrends(synch_stab_mod, pairwise ~ habitat, var = "synchrony") # fringing and backreef not different
# contrast                    estimate    SE  df z.ratio p.value
# Fringing - Backreef            0.181 0.107 Inf   1.695  0.3262
# Fringing - Forereef 10m        1.492 0.142 Inf  10.487  <.0001
# Fringing - Forereef 17m        2.436 0.149 Inf  16.336  <.0001
# Backreef - Forereef 10m        1.311 0.157 Inf   8.346  <.0001
# Backreef - Forereef 17m        2.255 0.161 Inf  14.008  <.0001
# Forereef 10m - Forereef 17m    0.944 0.181 Inf   5.222  <.0001
```

@@

#                        Chisq Df Pr(>Chisq)    
# richness_mean         17.645  1  2.662e-05 ***
# habitat               32.229  3  4.682e-07 ***
# richness_mean:habitat  8.075  3    0.04449 *  
hist(residuals(rich_stab_mod_site)) # blocky but fine
plot(residuals(rich_stab_mod_site) ~ fitted(rich_stab_mod_site)) # heteroskedastic - bring to group
r.squaredGLMM(rich_stab_mod_site) # 0.671
performance::r2(rich_stab_mod_site)
emtrends(rich_stab_mod_site, pairwise ~ habitat, var = "richness_mean") # no differences across habitats
# contrast                    estimate    SE  df z.ratio p.value
# Fringing - Backreef           0.1854 0.120 Inf   1.539  0.4138
# Fringing - Forereef 10m      -0.2155 0.145 Inf  -1.484  0.4471
# Fringing - Forereef 17m      -0.2876 0.214 Inf  -1.344  0.5349
# Backreef - Forereef 10m      -0.4009 0.165 Inf  -2.432  0.0711
# Backreef - Forereef 17m      -0.4730 0.228 Inf  -2.077  0.1607
# Forereef 10m - Forereef 17m  -0.0721 0.242 Inf  -0.298  0.9908
```

@@

hist(residuals(rich_stab_mod_site_fg)) # blocky but fine
plot(residuals(rich_stab_mod_site_fg) ~ fitted(rich_stab_mod_site_fg)) # heteroskedastic
r.squaredGLMM(rich_stab_mod_site_fg) # 0.563
performance::r2(rich_stab_mod_site_fg) # 0.551
emtrends(rich_stab_mod_site_fg, pairwise ~ habitat, var = "functional_richness_mean") 
# Fringing different from outer 10
# contrast                    estimate    SE  df z.ratio p.value
# Fringing - Backreef           0.2552 0.254 Inf   1.005  0.7463
# Fringing - Forereef 10m       1.2611 0.469 Inf   2.686  0.0364
# Fringing - Forereef 17m       0.3543 0.287 Inf   1.235  0.6045
# Backreef - Forereef 10m       1.0059 0.492 Inf   2.044  0.1720
# Backreef - Forereef 17m       0.0991 0.323 Inf   0.307  0.9900
# Forereef 10m - Forereef 17m  -0.9068 0.510 Inf  -1.778  0.2838
```

@@

### model
```{r synchrony staability mod site}
synch_stab_mod_site <- glmmTMB(cover_stability ~ synchrony*habitat, family = Gamma("log"), data = diversity_stability_synchrony_site)
summary(synch_stab_mod_site)
Anova(synch_stab_mod_site)
#                     Chisq Df Pr(>Chisq)    
# synchrony         22.0686  1  2.631e-06 ***
# habitat            9.3333  3    0.02517 *  
# synchrony:habitat 10.5561  3    0.01439 *  
hist(residuals(synch_stab_mod_site)) # almost uniform
plot(residuals(synch_stab_mod_site) ~ fitted(synch_stab_mod_site)) # same issue as always
r.squaredGLMM(synch_stab_mod_site) # 0.678
performance::r2(synch_stab_mod_site) # 0.702
emtrends(synch_stab_mod_site, pairwise ~ habitat, var = "synchrony") # backreef is different from forereef 17, fringing is a p = 0.05
#
# contrast                    estimate    SE  df z.ratio p.value
# Fringing - Backreef           -0.256 0.971 Inf  -0.263  0.9936
# Fringing - Forereef 10m        1.324 0.916 Inf   1.445  0.4709
# Fringing - Forereef 17m        2.930 1.142 Inf   2.567  0.0503
# Backreef - Forereef 10m        1.580 0.858 Inf   1.842  0.2534
# Backreef - Forereef 17m        3.186 1.095 Inf   2.909  0.0190
# Forereef 10m - Forereef 17m    1.606 1.047 Inf   1.534  0.4172
```

@@

### spatial synchrony
```{r model spatial synchrony}
spatial_synchrony_mod <- glmmTMB(cover_stability ~ spatial_synchrony*habitat, family = Gamma("log"), data = dss_spatial_2)
hist(residuals(spatial_synchrony_mod)) # a little skew
plot(residuals(spatial_synchrony_mod) ~ predict(spatial_synchrony_mod))
summary(spatial_synchrony_mod)
Anova(spatial_synchrony_mod)
#                             Chisq Df Pr(>Chisq)    
# spatial_synchrony         16.2914  1  5.431e-05 ***
# habitat                   19.5303  3  0.0002124 ***
# spatial_synchrony:habitat  6.7239  3  0.0812368 .  
performance::r2(spatial_synchrony_mod) # 0.64
emtrends(spatial_synchrony_mod, pairwise ~ habitat, var = "spatial_synchrony") # no sig differences but close
# contrast                    estimate    SE  df z.ratio p.value
# Fringing - Backreef            0.404 1.765 Inf   0.229  0.9958
# Fringing - Forereef 10m        1.365 1.754 Inf   0.778  0.8643
# Fringing - Forereef 17m        3.112 1.882 Inf   1.654  0.3484
# Backreef - Forereef 10m        0.961 0.857 Inf   1.121  0.6766
# Backreef - Forereef 17m        2.708 1.095 Inf   2.472  0.0644
# Forereef 10m - Forereef 17m    1.747 1.078 Inf   1.621  0.3668
```

@@

### mean synchrony
```{r mean synchrony}
mean_synchrony_mod <- glmmTMB(cover_stability ~ synchrony_mean*habitat, family = Gamma("log"), data = dss_spatial_2)
hist(residuals(mean_synchrony_mod)) # chunky
plot(residuals(mean_synchrony_mod) ~ predict(mean_synchrony_mod)) # fine
summary(mean_synchrony_mod)
Anova(mean_synchrony_mod)
#                         Chisq Df Pr(>Chisq)    
# synchrony_mean         32.914  1  9.634e-09 ***
# habitat                18.523  3  0.0003430 ***
# synchrony_mean:habitat 19.062  3  0.0002654 ***
performance::r2(mean_synchrony_mod)
emtrends(mean_synchrony_mod, pairwise ~ habitat, var = "synchrony_mean") # no sig differences but close
# contrast                    estimate    SE  df z.ratio p.value
# Fringing - Backreef            -3.26 1.503 Inf  -2.166  0.1327
# Fringing - Forereef 10m         2.22 0.995 Inf   2.231  0.1149
# Fringing - Forereef 17m         3.37 1.220 Inf   2.764  0.0291
# Backreef - Forereef 10m         5.48 1.622 Inf   3.377  0.0041
# Backreef - Forereef 17m         6.63 1.769 Inf   3.748  0.0010
# Forereef 10m - Forereef 17m     1.15 1.363 Inf   0.846  0.8322
```

@@

### mean stability plot
```{r alpha gamma stability mod plot}
alpha_gamma_stab_mod <- glmmTMB(cover_stability ~ stability_mean*habitat, family = Gamma("log"), data = dss_spatial_2)
summary(alpha_gamma_stab_mod)
Anova(alpha_gamma_stab_mod)
#                          Chisq Df Pr(>Chisq)    
# stability_mean         61.2725  1   4.97e-15 ***
# habitat                12.9323  3   0.004785 ** 
# stability_mean:habitat  9.4832  3   0.023511 *  
performance::r2(alpha_gamma_stab_mod)
